Wall-to-wall mapping of peat depth from Lidar terrain and airborne radiometrics in Norwegian landscapes

### 2024-10-30 ms/ms.Rmd
table style for clarity

We created the same suite of peat depth predictors for both sites (25 continuous and 1 categorical; Table \@ref(tab:preds)).
All continuous predictors were derived either from an airborne radiometric survey or from a DTM.
From the radiometric surveys we simply used the four variables produced by the surveyors (Geological Survey of Norway): ground concentration of Potassium, Thorium, Uranium, as well as total count.
From the DTMs we calculated several land surface parameters, ranging from simple terrain indices to more complex geomorphometric and hydrological variables [@maxwellLandsurfaceParametersSpatial2022].
The categorical predictor was peat depth class, from a national map dataset.
Complete descriptions of predictors follow below.
```{r preds, echo=FALSE, message=FALSE}
df <- readr::read_csv("tables/predictors.csv")
knitr::kable(df, booktabs = FALSE, caption = "Candidate predictors of peat depth.")
```

@@

We calculated partial dependence with the *pdp* R package [v.0.8.1, @greenwellPdpPackageConstructing2017].
For the top six most important variables, we plotted both partial dependence and individual conditional expectation (ICE), to show the average effect of the predictor on the outcome and the variation in the effect across observations, repectively [@goldsteinPeekingBlackBox2015].
Non-parallel ICE lines indicate the presence of interactions between predictors.

>>>>

We created the same suite of peat depth predictors for both sites (25 continuous and 1 categorical; Table \@ref(tab:preds)).
All continuous predictors were derived either from an airborne radiometric survey or from a DTM.
From the radiometric surveys we simply used the four variables produced by the surveyors (Geological Survey of Norway): ground concentration of Potassium, Thorium, Uranium, as well as total count.
From the DTMs we calculated several land surface parameters, ranging from simple terrain indices to more complex geomorphometric and hydrological variables [@maxwellLandsurfaceParametersSpatial2022].
The categorical predictor was peat depth class, from a national map dataset.
Complete descriptions of predictors follow below.
```{r preds, echo=FALSE, message=FALSE}
df <- readr::read_csv("tables/predictors.csv")
knitr::kable(df, booktabs = FALSE, vline = "", linesep = "",
             col.names = c("Name", "Description"),
             caption = "Candidate predictors of peat depth.")
```

@@

We calculated partial dependence with the *pdp* R package [v.0.8.1, @greenwellPdpPackageConstructing2017].
For the top six most important variables, we plotted both partial dependence and individual conditional expectation (ICE), to show the average effect of the predictor on the outcome and the variation in the effect across observations, respectively [@goldsteinPeekingBlackBox2015].
Non-parallel ICE lines indicate the presence of interactions between predictors.

### 2024-10-30 ms/ms.tex
table style for clarity

\caption{\label{tab:preds}Candidate predictors of peat depth.}
\centering
\begin{tabular}[t]{l|l}
\hline
name & description\\
\hline
radK & Potassium ground concentration\\
\hline
radTh & Thorium ground concentration\\
\hline
radU & Uranium ground concentration\\
\hline
radTC & Total count of gamma radiation\\
\hline
elevation & Mean elevation\\
\hline
slope1m & Mean of 1 m slope\\
\hline
TPI1m & Mean of 1 m topographic position index\\
\hline
TRI1m & Mean of 1 m terrain ruggedness index\\
\hline
roughness1m & Mean of 1 m roughness\\
\hline
slope10m & 10 m slope\\
\hline
TPI10m & 10 m topographic position index\\
\hline
TRI10m & 10 m terrain ruggedness index\\
\hline
roughness10m & 10 m roughness\\
\hline
MRVBF & Multi-resolution valley bottom flatness\\
\hline
TWI5m & Mean of 5 m topographic wetness index\\
\hline
TWI10m & 10 m topographic wetness index\\
\hline
TWI20m & Bilinear interpolation of 20 m topographic wetness index\\
\hline
TWI50m & Bilinear interpolation of 50 m topographic wetness index\\
\hline
DTW2500 & Depth-to-water index, flow initiation area of 0.25 ha\\
\hline
DTW5000 & Depth-to-water index, flow initiation area of 0.5 ha\\
\hline
DTW10000 & Depth-to-water index, flow initiation area of 1 ha\\
\hline
DTW20000 & Depth-to-water index, flow initiation area of 2 ha\\
\hline
DTW40000 & Depth-to-water index, flow initiation area of 4 ha\\
\hline
DTW80000 & Depth-to-water index, flow initiation area of 8 ha\\
\hline
DTW160000 & Depth-to-water index, flow initiation area of 16 ha\\
\hline
DMK & DMK peat depth class, categorical with 3 levels\\
\hline
\end{tabular}
\end{table}

@@

We calculated partial dependence with the \emph{pdp} R package \citep[v.0.8.1,][]{greenwellPdpPackageConstructing2017}.
For the top six most important variables, we plotted both partial dependence and individual conditional expectation (ICE), to show the average effect of the predictor on the outcome and the variation in the effect across observations, repectively \citep{goldsteinPeekingBlackBox2015}.
Non-parallel ICE lines indicate the presence of interactions between predictors.

>>>>

\caption{\label{tab:preds}Candidate predictors of peat depth.}
\centering
\begin{tabular}[t]{ll}
\hline
Name & Description\\
\hline
radK & Potassium ground concentration\\
radTh & Thorium ground concentration\\
radU & Uranium ground concentration\\
radTC & Total count of gamma radiation\\
elevation & Mean elevation\\
slope1m & Mean of 1 m slope\\
TPI1m & Mean of 1 m topographic position index\\
TRI1m & Mean of 1 m terrain ruggedness index\\
roughness1m & Mean of 1 m roughness\\
slope10m & 10 m slope\\
TPI10m & 10 m topographic position index\\
TRI10m & 10 m terrain ruggedness index\\
roughness10m & 10 m roughness\\
MRVBF & Multi-resolution valley bottom flatness\\
TWI5m & Mean of 5 m topographic wetness index\\
TWI10m & 10 m topographic wetness index\\
TWI20m & Bilinear interpolation of 20 m topographic wetness index\\
TWI50m & Bilinear interpolation of 50 m topographic wetness index\\
DTW2500 & Depth-to-water index, flow initiation area of 0.25 ha\\
DTW5000 & Depth-to-water index, flow initiation area of 0.5 ha\\
DTW10000 & Depth-to-water index, flow initiation area of 1 ha\\
DTW20000 & Depth-to-water index, flow initiation area of 2 ha\\
DTW40000 & Depth-to-water index, flow initiation area of 4 ha\\
DTW80000 & Depth-to-water index, flow initiation area of 8 ha\\
DTW160000 & Depth-to-water index, flow initiation area of 16 ha\\
DMK & DMK peat depth class, categorical with 3 levels\\
\hline
\end{tabular}
\end{table}

@@

We calculated partial dependence with the \emph{pdp} R package \citep[v.0.8.1,][]{greenwellPdpPackageConstructing2017}.
For the top six most important variables, we plotted both partial dependence and individual conditional expectation (ICE), to show the average effect of the predictor on the outcome and the variation in the effect across observations, respectively \citep{goldsteinPeekingBlackBox2015}.
Non-parallel ICE lines indicate the presence of interactions between predictors.
